Modeling browser-based distributed evolutionary computation systems

### 2015-03-21 modeling-volunteer-computing-gecco.tex
De-nonymizes and corrects for ArXiV waiting for @fergunet corrections

\documentclass{sig-alternate}

@@

\keywords{Evolutionary computation, volunteer computing, distributed
  computing, node.js, javascript}

@@

\begin{figure*}[h!tb]
  \centering
  \includegraphics[trim=0cm 2cm 0cm 3cm, width=10cm]{img/screenshot.png}
  \label{fig:sshot}
  \caption{Screenshot of the talk that includes at the graph
    that shows the progression of the evolutionary
    algorithm at the top right corner}
\end{figure*}
%FERGU: No puedes poner el gráfico un poco más grande? (y de paso 
% recortar mejor la imagen para que no se vea el texto de la derecha)
In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences\footnote{Names withheld for the double blind review} and
it has the appearance shown in Figure \ref{fig:sshot}. During the
conference, it was revealed to the users that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. As it can be seen in the
image, the upper right corner shows the progress of the evolution. It
stops when the solution has been reached.

Above the experiment, from the point of view of the user, has been
described. What it is actually running is an island model
\cite{muhlenbein1991parallel} in every browser that uses the server as
a {\em shuttle} to transfer individuals from one browser to another in
what can be eventually a fully connected topology. 
%FERGU: explica esto: los navegadores se conectan entre sí o sólo a tu servidor?
This deals with
several problems. The connection is stateless: islands connect to the
server to send and receive a single individual but there is no task
balancing: all islands start clicking on an URL and finish when they
browse off to the next page. Fault tolerance is implicit, in the sense
that no island has information that cannot be, or generated, somewhere
else, although if the server fails the experiment results might be
lost. All operation is asynchronous, with islands entering and leaving
the experiment all of their own.

@@

This fitness is implemented as part of a (name hidden) evolutionary
%FERGU: creo que si usas un framework libre puedes dejarlo visible
algorithm library written in node.js, a server-based implementation of
JavaScript. However, since it is going to be run on the browser, a
procedure called ``browserification'' must be applied to it so that it
can run, without modifications, in any browser. The main difference
between JavaScript on the browser and node.js is the way they load the
%FERGU: puesto .js para distinguir la palabra node del nombre de la librería
external libraries. The algorithm used a population of 256
individuals, random initialization, a canonical evolutionary algorithm
with an elite of 2 individuals and crossover and then mutation applied
to all the population pool, that was generated using 2-individual
tournament selection. Once again, the particular evolutionary
algorithm used is not so important, since we were rather focused on
the user behavior, although it is quite clear that the selective
pressure of the algorithm will have an impact in its performance and
will play for or against its asynchrony in many different and complex
ways \cite{jj:2008:PPSN}. %FERGU: este párrafo me lo copio para el de la ASOCO ;)

The initial code was only modified to represent the fitness in a chart
(that uses the {\tt chart.js} JavaScript library) so that the user has
a visual feedback of what is going on. The {\em migration} part is
also added: after every 100 generations, a single individual is sent
to the server and another request is made to the server of a single
and random individual. This makes the procedure less greedy, and also
decreases the probability of getting the same individual that was sent
some generations back. These two operations are made in {\em fire and
  forget} style without any error checking, so that if the server is
  %FERGU: Qué es fire and forget? (se puede intuir, pero dejarlo claro)
down, or indeed if the algorithm is run locally, it is not interrupted
and continues operation until the solution is found, a string with all
ones.

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in a GitHub
repository (address hidden for double blind review). These results
will be analyzed in the next section.

@@

This paper is supported in part by several grants hidden for
anonymity.

\bibliographystyle{abbrv}
\bibliography{geneura,volunteer,javascript,ror-js,GA-general}

>>>>

\documentclass{article}

@@

@@

In order to test the volunteer computing environment, a presentation
describing a low cost volunteer computing environment was
created. This presentation was actually delivered in several
conferences\footnote{Names withheld for the double blind review}. During the
conference, it was revealed to the persons attending it that they were participating
in the experiment. The same procedure was used when trying to gather
users in social networks: a description of the container (the
presentation) and disclosure of its purpose. The upper right corner of
the presentation shows the progress of the evolution. It
stops when the solution has been reached.

Above the experiment, from the point of view of the user, has been
described. What it is actually running is an island model
\cite{muhlenbein1991parallel} in every browser that uses the server as
a {\em shuttle} to transfer individuals from one browser to another in
what can be eventually a fully connected topology; however, all
connections take place between the browser and the server.
This deals with
several problems. The connection is stateless: islands connect to the
server to send and receive a single individual but there is no task
balancing: all islands start clicking on an URL and finish when they
browse off to the next page. Fault tolerance is implicit, in the sense
that no island has information that cannot be, or generated, somewhere
else, although if the server fails the experiment results might be
lost. All operation is asynchronous, with islands entering and leaving
the experiment all of their own.

@@

This fitness is implemented as part of the Nodeo evolutionary
algorithm library written in node.js, a server-based implementation of
JavaScript. However, since it is going to be run on the browser, a
procedure called ``browserification'' must be applied to it so that it
can run, without modifications, in any browser. The main difference
between JavaScript on the browser and node.js is the way they load the
external libraries. The algorithm used a population of 256
individuals, random initialization, a canonical evolutionary algorithm
with an elite of 2 individuals and crossover and then mutation applied
to all the population pool, that was generated using 2-individual
tournament selection. Once again, the particular evolutionary
algorithm used is not so important, since we were rather focused on
the user behavior, although it is quite clear that the selective
pressure of the algorithm will have an impact in its performance and
will play for or against its asynchrony in many different and complex
ways \cite{jj:2008:PPSN}.

The initial code was only modified to represent the fitness in a chart
(that uses the {\tt chart.js} JavaScript library) so that the user has
a visual feedback of what is going on. The {\em migration} part is
also added: after every 100 generations, a single individual is sent
to the server and another request is made to the server of a single
and random individual. This makes the procedure less greedy, and also
decreases the probability of getting the same individual that was sent
some generations back. These two operations are made in {\em fire and
  forget} style, that is, without any error checking, so that if the server is
down, or indeed if the algorithm is run locally, it is not interrupted
and continues operation until the solution is found, a string with all
ones.

@@

As indicated in the introduction, every part of the experiment and
data gathered have been released as free software in the GitHub
repository \url{https://github.com/JJ/idc-keynote}. These results
will be analyzed in the next section.

@@

This work has been supported in part by TIN2011-28627-C04-02 and
TIN2014-56494-C4-3-P (Spanish Ministry of Economy and Competitivity),
SPIP2014-01437 (Direcci{\'o}n General de Tr{\'a}fico) and PYR-2014-17
GENIL project (CEI-BIOTIC Granada).

\bibliographystyle{abbrv}
\bibliography{geneura-latin1,volunteer,javascript,ror-js,GA-general}
